### 2021-06-27 paper.tex
changed v2 <-> a2 (+ other minor changes) + organization added

\section{Introduction and summary}
\label{sec:introduction}
Our understanding of the concept of symmetries in quantum field theories (QFT) has been greatly improved in the last several years.
We now have the concept of $p$-form symmetries acting on $p$-dimensional objects \cite{Gaiotto:2014kfa}.
This concept  gives a unifying point of view to both
ordinary symmetries acting on point operators for $p=0$
and center symmetries of gauge theories acting on Wilson line operators for $p=1$.
In addition, the 't Hooft magnetic flux \cite{tHooft:1979rtg} can now be thought of as a background gauge field for the 1-form center symmetry.
It is also realized more recently that 0-form symmetries and 1-form symmetries can not only coexist in a direct product but also mix in a more intricate manner.
They can have mixed anomalies between them,
or they can also combine to form a symmetry structure called 2-groups \cite{Cordova:2018cvg,Benini:2018reh}.

In this paper we study these issues in the case of 4d $\so(N_c)$ gauge theories with 
$N_f$ flavors of fermion fields in vector representation.
Let us quickly recall the 0-form and 1-form symmetries these theories have.

As for the 1-form symmetry, we first need to recall that 
such theories come in three versions, $\Spin$, $\SO_+$ and $\SO_-$,
distinguished by the global form of the gauge group ($\Spin$ vs.~$\SO$)
and further by the choice of a discrete theta angle ($\SO_+$ vs.~$\SO_-$) \cite{Aharony:2013hda}\footnote{%
This is when the theories are considered on spin manifolds.
On more general manifolds, a further distinction needs to be made \cite{Ang:2019txy}.
For simplicity, we only consider spin manifolds in this paper.
}.
They also differ by the nontrivial line operator they possess: 
the $\Spin$ theory has the Wilson line $W$ in the spinor representation,
the $\SO_+$ theory has the 't Hooft line $H$ which is mutually non-local with respect to $W$,
and the $\SO_-$ theory has the dyonic line $D=WH$. 
Furthermore, these line operators are charged under corresponding $\bZ_2$ 1-form symmetries,
which we respectively call electric, magnetic and dyonic 1-form symmetries.

@@

Take for example the $Spin(2n_c)$ gauge theory with $2n_f$ flavors,
when $n_c$ is odd.
Take two copies of  the Wilson line $W$ in the spinor representation. 
They form a Wilson line in the vector representation.
This can be screened by a dynamical fermion, which was why $W^2=1$ 
as far as the 1-form symmetry charge was concerned.
Now let us recall that this dynamical fermion transforms nontrivially under $-1\in \SU(2n_f)$.
Therefore, when we take the flavor symmetry into account, $W^2$ is still nontrivial.
As we will recall below,  formally this means that the $\bZ_2$ 1-form symmetry extends the $\SU(2n_f)/\bZ_2$ 0-form symmetry in a nontrivial manner, forming a nontrivial 2-group  $H$\begin{equation}
0\to \bZ_2[1] \to H\to \SU(2n_f)/\bZ_2\to 0,\label{2-group}
\end{equation}
whose Postnikov class is specified by \begin{equation}
\beta v_2 \in H^3(B(\SU(2n_f)/\bZ_2);\bZ_2),\label{postnikov}
\end{equation}
where $v_2$ is the generator of $H^2(\SU(2n_f)/\bZ_2;\bZ_2)=\bZ_2$ and $\beta$ is the Bockstein.

The $\SO_+$ gauge theory is then obtained by gauging the $\bZ_2$ 1-form symmetry \cite{Kapustin:2014gua}.
The presence of the fermions significantly complicates the analysis.
For the moment let us suppose that we have $N_f$ scalars instead of fermions in the vector representation.
Then, the argument of \cite{Tachikawa:2017gyf} immediately applies, and 
we see that the $\bZ_2$ 1-form symmetry of the $\SO(2n_f)$ theory and the $\SU(2n_f)/\bZ_2$ 0-form flavor symmetry remains a direct product but with a mixed anomaly given by \begin{equation}
2\pi i \cdot \frac12 \int B\beta v_2.\label{mixed}
\end{equation}

@@

The rest of the paper is organized as follows ...

\section{2-group structure}
\label{sec:2-group}
Let us first study whether the $\bZ_2$ 1-form symmetry and the $\SU(2n_f)/\bZ_2$ flavor symmetry form a nontrivial 2-group or not. 
This can be found rather physically by studying the line operators.

\subsection{$\Spin$}
We start by discussing the $\Spin(2n_c)$ gauge theories. 
We first recall that the center of $\Spin(2n_c)$ is $\bZ_2\times \bZ_2$ when $n_c$ is even and $\bZ_4$ when $n_c$ is odd.
This corresponds to the fact that the tensor square of a spinor representation contains the identity representation when $n_c$ is even while it contains the vector representation when $n_c$ is odd.

We now consider the Wilson line $W$ in the spinor representation in the $\Spin(2n_c)$ gauge theory with $2n_f$ fermions in the vector representation.
When $n_c$ is even, $W^2$ contains the identity representation, and therefore we simply have a $\bZ_2$ 1-form symmetry independent of the $\su$ flavor symmetry, and there is nothing to see here.

When $n_c$ is odd, $W^2$ contains the vector representation.
This can be screened by the dynamical fermion, which however carries the fundamental representation of $\su(2n_f)$ flavor symmetry, 
and in particular transforms nontrivially under $-1\in \SU(2n_f)$.
In other words, the flavor Wilson line in the vector representation of $\SU(2n_f)$ can now be considered as the square  of the gauge Wilson line in the spinor representation of $\Spin(2n_c)$.
This means that we have the following extension of groups \begin{equation}
0\to \underbrace{\bZ_2}_{\substack{\text{subgroup of}\\
\text{rep.~of center of $\SU(2n_f)$}}}
\to \bZ_4 
\to \underbrace{\bZ_2}_{\substack{\text{group of gauge Wilson lines }\\
\text{up to screening}}} \to 0.
\label{charge-extension}
\end{equation}

As the groups of charges of $\su(2n_f)$ 0-form symmetry and $\bZ_2$ 1-form symmetry are combined nontrivially, 
the 0-form symmetry group and the 1-form symmetry group are also combined nontrivially.
This can be seen most clearly by considering background fields for the symmetry groups.

The fermion fields are in the vector of $\SO(2n_c)$ and in the fundamental of $\SU(2n_f)$, and therefore is a representation of $G=[\SO(2n_c)\times \SU(2n_f)]/\bZ_2$.
Given a $G$-bundle on a manifold $X$,
there is an $\SO(2n_c)/\bZ_2$ bundle and an $\SU(2n_f)/\bZ_2$ bundle associated to it.
Let us denote by $a_2,v_2\in H^2(X;\bZ_2)$ the obstruction classes controlling whether they lift to $\SO(2n_c)$ and $\SU(2n_f)$ respectively. 
Then we have $a_2=v_2$ for a $G$-bundle.
The flavor Wilson line in the fundamental representation is charged under $-1\in \SU(2n_f)$ in the center,
and $v_2$ can be considered as the background field for this $\bZ_2$ 1-form center symmetry.

Now, without the flavor background,  the background $E\in H^2(X;\bZ_2)$ for the electric $\bZ_2$ one-form symmetry of the $\Spin(2n_c)$ theory sets the Stiefel-Whitney class $w_2\in H^2(X;\bZ_2)$ of the $\SO(2n_c)$ gauge bundle to be $E=w_2$, which controls whether it lifts to a $\Spin(2n_c)$ bundle.
When the flavor background $v_2$ is nontrivial,
the obstruction class $a_2$ controlling the lift from $\SO(2n_c)/\bZ_2$ to $\SO(2n_c)$ is nontrivial.
In this situation when $n_c$ is odd, $w_2$ can no longer be defined as a closed cochain; rather it satisfies $\delta w_2 = \beta a_2$, where $\beta$ is the Bockstein operation,
since together they specify the obstruction class $x_2\in H^2(X;\bZ_4)$ 
controlling the lift from $\SO(2n_c)/\bZ_2=\Spin(2n_c)/\bZ_4$ to $\Spin(2n_c)$.
As $E=w_2$ and $a_2=v_2$, we conclude that the background fields satisfy \begin{equation}
\delta E=\beta v_2.
\label{ba}
\end{equation}
This means that the $\bZ_2$ 1-form symmetry and the $\SU(2n_f)/\bZ_2$ 0-form flavor symmetry form the 2-group $H$ fitting in the sequence \begin{equation}
0\to \bZ_2[1]\to H \to \SU(2n_f)/\bZ_2 \to 0
\label{group-extension}
\end{equation} whose Postnikov class is $\beta v_2 \in H^3(B(\SU(2n_f)/\bZ_2);\bZ_2)$.

Before proceeding, we note that having the extension of groups of charges of line operators as in \eqref{charge-extension}
is equivalent to having a nontrivial 2-group extension \eqref{group-extension} 
whose background fields satisfy \eqref{ba}.
Therefore, to find a nontrivial 2-group extension, we can simply study the group of charges of line operators,
which we will carry out for $\SO_\pm$ gauge theories next.

@@

We would like to study how the magnetic $\bZ_2$ 1-form symmetry of the $\SO(2n_c)_+$ theory is combined with the $\su(n_f)$ flavor symmetry.
According to the discussions in the previous subsection, 
we consider what happens if we take two copies of the 't Hooft line operator $H$ and fuse them.
At the very naive level, $H^2$ can be screened by a dynamical monopole,
but dynamical monopoles can receive flavor and gauge center charges from the fermion zero modes.
To study these issues, it is useful to deform the theories to make them simpler.

For this purpose, we perform the following steps:
\begin{itemize}
\item We reduce the flavor symmetry from $\su(2n_f)$ to $\usp(2n_f)$. 
The fundamental representation still transforms nontrivially under $-1\in \USp(2n_f)$, which is enough for our purposes.
\item We add an adjoint scalar $\Phi_{[ab]}$ and the interaction $\psi^{ai}_\alpha \psi^{bj}_\beta J_{ij} \Phi_{ab}\epsilon^{\alpha\beta} +cc$, where $J_{[ij]}$ is the constant matrix for the $\usp(2n_f)$ part.
\item We give a vev to $\Phi_{ab}$ to break $\SO(2n_c)$ to $\SO(2)^{n_c}$.
\end{itemize}

The 't Hooft lines in the resulting $\SO(2)^{n_c}$ theory can be labeled by their magnetic charges $(m_1,\ldots,m_{n_c})\in \bZ^{n_c}$.
The dynamical monopoles have the charges in the `adjoint class', which are
in the root lattice $\Lambda$ of $\SO(2n_c)$.

The group of the magnetic charges of 't Hooft lines up to screening by the dynamical monopoles is then \begin{equation}
\bZ^{n_c}/\Lambda = \bZ_2,
\end{equation}
which agrees with the 1-form symmetry before the deformation.
We now would like to study how this $\bZ_2$ is combined with the flavor/gauge center $\bZ_2$ charge.

@@

Let us say we chose the standard $\phi$ such that the simple roots are $(1,-1,\ldots,0)$,
$(0,1,-1,\ldots,0)$, \ldots, $(0,\ldots,1,-1)$ and $(0,\ldots,1,+1)$,
which we call simple dynamical monopoles.

We now consider the group $\bZ^{n_c}\times \bZ_2$ which combine the magnetic charges $\bZ^{n_c}$ and the flavor/gauge center charge $\bZ_2$.
What we are after is the quotient of $\bZ^{n_c}\times \bZ_2$ 
by the subgroup generated by the charges of simple dynamical monopoles,
which we denote by
$(1,-1,\ldots, 0; q_1)$, $(0,1,-1,\ldots, 0; q_2)$, \ldots,
$(0,\ldots, 1,-1; q_{n_c-1})$ and $(0,\ldots,1,+1; q_{n_c})$, respectively.

Let us determine this quotient.
We do not have to determine $q_1$ to $q_{n_c-1}$. 
We simply use them to rewrite any charge vector $
(m_1,\ldots, m_{n_c-2}, m_{n_c-1},m_{n_c}; q)
$
into the form $
(0,\ldots,0, m, m' ; q').
$
We then have to determine $q_{n_c-1}$ and $q_{n_c}$.

This reduces the study to the case of $\so(4)\simeq \su(2)_1 \times \su(2)_2$.
The monopoles associated to the simple roots are just 't Hooft-Polyakov monopoles associated to the two factors of $\su(2)$'s.
The vev of the adjoint scalar in this basis can be written as $(a_1,a_2)$, which we assume $a_1>a_2>0$.
The fermion is in the vector representation of $SO(4)$.
Under the monopole in $\su(2)_1$, it is a doublet coupled to an adjoint vev of size $a_1$ with bare mass $a_2$,
and similarly for the monopole in $\su(2)_2$.

Now, the explicit analysis in \cite[Sec.~IV]{Callias:1977kg} concerning the number of zero modes in the 't Hooft-Polyakov monopole says that 
a doublet coupled to an adjoint vev of size $a$ with bare mass $\mu$ has
 a zero mode if $|a|>|\mu|$ and 
 it has no zero mode if $|a|<|\mu|$.

With our assumption $a_1>a_2>0$, this means that the monopole in $\su(2)_1$ has a zero mode while the one in $\su(2)_2$ does not have any.
In our original basis, this means that the monopole with $(0,\ldots,1,-1;q_{n_c-1})$ does not have any zero modes and $q_{n_c-1}=0$,
while the one with $(0,\ldots,1,+1;q_{n_c})$  (which is in $\su(2)_1$) has two zero modes per flavor.

The one-form symmetry group is obtained by dividing $\bZ^2\times \bZ_2$ by the subgroup generated by $(1,-1;0)$ and $(1,+1;q_{n_c})$.
This is $\bZ_2\times \bZ_2$ or $\bZ_4$ depending on whether $q_{n_c}$ is $0$ or $1$.

Let us determine $q_{n_c}$.
We saw two zero modes per flavor; 
this means that there are fermionic zero modes transforming in \begin{equation}
V_{2} \otimes R_{2n_f} 
\end{equation} where $V_2$ is the doublet of $\su(2)_2$ (which is actually broken to $\u(1)$ but keeping $\su(2)_2$ representation is useful in organizing the answer), 
$R_{2n_f}$ is the fundamental of $\usp(2n_f)$,
and we need to impose the reality condition using the pseudoreality of both factors,
so that there are $4n_f$ Majorana fermion in total.

To determine the flavor/gauge center charge of the monopole,
it suffices to consider the case $n_f=1$;
the general case is given by simply multiplying it by $n_f$.
When $n_f=1$, there are $4$ Majorana fermions.
Quantizing them, we find the monopoles in \begin{equation}
V_2 \otimes \mathbf{1} \oplus \mathbf{1}\otimes R_2.
\end{equation}
It has the `vector' charge under $\usp(2)$ flavor symmetry or is a doublet under $\su(2)_2$,
which corresponds to the `vector' charge under $\so(4)$ gauge symmetry.
In either case, they have the flavor/gauge center charge $1 \in \{0,1\}=\bZ_2$.
Therefore we conclude the flavor/gauge center charge $q_{n_c}$ is simply given by $n_f$ mod 2.

@@

The analysis of the $\SO(2n_c)_-$ theory is largely the same;
the only difference is that the discrete theta angle gives an additional gauge center charge to the simple dynamical monopole with the magnetic charge $(0,0,\ldots,1,1)$, so that $q_{n_c}=n_f+n_c$ mod 2.
Therefore, we conclude the following:
for the $\SO(2n_c)_-$ theory,
the group $\bZ_2$ of magnetic charges of 't Hooft lines is extended by the flavor/gauge center symmetry $\bZ_2$ to become $\bZ_4$ when $n_f+n_c$ is odd,
while they remain separate when $n_f+n_c$ is even.

@@

\begin{table}
\centering
\begin{tabular}{r@{\vphantom{$\Bigm|$}\,}l|ccc}
$(n_c,$&$n_f)$ & $\Spin$ & $\SO_+$ & $\SO_-$\\
\hline
(even,&even) & product & product & product \\
(odd,&even) & extended & product & extended \\
(even,&odd) & product & extended & extended \\
(odd,&odd) & extended & extended & product
\end{tabular}
\caption{How the $\bZ_2$ 1-form symmetry and the flavor symmetry $\SU(N_f)/\bZ_2$ are combined
in $\so(2n_c)$ QCD.
`product' means that they form a direct product,
and `extended' means that they form a nontrivial 2-group.
\label{table:2group}}
\end{table}

@@

Let us say that  a four-dimensional spin QFT $Q$ with $\bZ_2$ 1-form symmetry is given.
We denote its partition function on a manifold $X$ by $Z_Q[E]$, 
where we suppress the dependence on $X$ in the notation and $E\in H^2(X;\bZ_2)$ is the background field for the 1-form symmetry.
We then define $SQ$ and $TQ$ to be QFTs with partition functions given by the formula \begin{equation}
Z_{SQ}[B] \propto \sum_{E} (-1)^{\int_X B\cup E} Z_Q{B},\qquad
Z_{TQ}[E]= (-1)^{\int_X \tfrac12 \fP(E)} Z_Q[E].
\end{equation}
We can show that $S^2=T^2=1$ and $(ST)^3=1$, meaning that they generate $\SL(2,\bZ_2)$.
This operation was considered in  \cite{Gaiotto:2014kfa} as an analogue of the $\SL(2,\bZ)$ action on 3d theories with $\U(1)$ symmetry of  \cite{Witten:2003ya} and then further studied in \cite{Bhardwaj:2020ymp}.

>>>>

\section{Introduction and summary}
\label{sec:introduction}
Our understanding of the concept of symmetries in quantum field theories (QFT) has been greatly improved in the last several years.
We now have the concept of $p$-form symmetries acting on $p$-dimensional objects \cite{Gaiotto:2014kfa}.
This concept  gives a unifying point of view to both
ordinary symmetries acting on point operators for $p=0$
and center symmetries of gauge theories acting on Wilson line operators for $p=1$.
In addition, the 't Hooft magnetic flux \cite{tHooft:1979rtg} can now be thought of as a background gauge field for the 1-form center symmetry.
It is also realized more recently that 0-form symmetries and 1-form symmetries can not only coexist in a direct product but also mix in a more intricate manner.
They can have mixed anomalies between them,
or they can also combine to form a \textit{higher} symmetry structure called 2-groups \cite{Cordova:2018cvg,Benini:2018reh}.

In this paper, we study these issues in the case of 4d $\so(N_c)$ gauge theories with 
$N_f$ flavors of fermion fields in the vector representation.
Let us quickly recall the 0-form and 1-form symmetries these theories have.

As for the 1-form symmetry, we first need to recall that 
such theories come in three versions, $\Spin$, $\SO_+$ and $\SO_-$,
distinguished firstly by the global form of the gauge group ($\Spin$ vs.~$\SO$)
and further by the choice of a discrete theta angle ($\SO_+$ vs.~$\SO_-$) \cite{Aharony:2013hda}\footnote{%
This is when the theories are considered on spin manifolds.
On more general manifolds, a further distinction needs to be made \cite{Ang:2019txy}.
For simplicity, we only consider spin manifolds in this paper.
}.
They also differ by the nontrivial line operator they possess: 
the $\Spin$ theory has the Wilson line $W$ in the spinor representation,
the $\SO_+$ theory has the 't Hooft line $H$ which is mutually non-local with respect to $W$,
and the $\SO_-$ theory has the dyonic line $D=WH$. 
Furthermore, these line operators are charged under corresponding $\bZ_2$ 1-form symmetries,
which we respectively call electric, magnetic and dyonic 1-form symmetries.

@@

Take for example the $\Spin(2n_c)$ gauge theory with $2n_f$ flavors.
When $n_c$ is odd,
two copies of the Wilson line $W$ in the spinor representation
form a Wilson line in the vector representation.
This can be screened by a dynamical fermion, and that was why $W^2=1$ 
as far as the 1-form symmetry charge was concerned.
Now let us recall that this dynamical fermion transforms nontrivially under $-1\in \SU(2n_f)$.
Therefore, when we further take the flavor symmetry into account, $W^2$ is still nontrivial.
As we will recall below,  formally this means that the $\bZ_2$ 1-form symmetry extends the $\SU(2n_f)/\bZ_2$ 0-form symmetry in a nontrivial manner, forming a 2-group $H$\begin{equation}
0\to \bZ_2[1] \to H\to \SU(2n_f)/\bZ_2\to 0,\label{2-group}
\end{equation}
whose Postnikov class is specified by \begin{equation}
\beta a_2 \in H^3(B(\SU(2n_f)/\bZ_2);\bZ_2),\label{postnikov}
\end{equation}
where $a_2$ is the generator of $H^2(B(\SU(2n_f)/\bZ_2);\bZ_2)=\bZ_2$~\cite{KonoMimura1974}
and $\beta$ is the Bockstein homomorphism.

The $\SO_+$ gauge theory is then obtained by gauging the $\bZ_2$ 1-form symmetry \cite{Kapustin:2014gua}.
The presence of the fermions significantly complicates the analysis.
For the moment let us suppose that we have $N_f$ scalars instead of fermions in the vector representation.
Then, the argument of \cite{Tachikawa:2017gyf} immediately applies, and 
we see that the $\bZ_2$ 1-form symmetry of the $\SO(2n_f)$ theory and the $\SU(2n_f)/\bZ_2$ 0-form flavor symmetry remains a direct product but with a mixed anomaly given by \begin{equation}
2\pi i \cdot \frac12 \int B\beta a_2.\label{mixed}
\end{equation}

@@

The rest of the paper is organized as follows.
In Sec.~\ref{sec:2-group},
we determine exactly when the electric/magnetic/dyonic $\bZ_2$ 1-form symmetries
and the $\SU(2n_f)/\bZ_2$ flavor 0-form symmetry are mixed nontrivially,
by examining the charges of line operators in each theory.
In Sec.~\ref{sec:sl2z},
we exploit the $\SL(2,\bZ_2)$ actions on $\so(2n_f)$ QCDs to determine the 't Hooft anomalies they possess.
Combining the results with those obtained in Sec.~\ref{sec:2-group},
one can completely determine the structures of symmetries and anomalies of $\so(2n_f)$ QCDs,
and can further confirm that they are indeed compatible with the Intriligator-Seiberg duality.
Although the result itself is satisfactory, the analysis leading to it is somewhat ad-hoc, 
so we partially complement it with the more direct computation of fermion anomalies in Sec.~\ref{sec:fermion}.
Finally, the two appendices provide technical details of the mathematical facts used in the main part;
in Appendix~\ref{sec:bordism}, we compute relevant bordism groups capturing the fermion anomalies associated to various symmetries;
and in Appendix~\ref{sec:nonclosedP}, we describe some subtleties concerning the Pontrjagin square.

\section{2-group structure}
\label{sec:2-group}
Let us first study whether the $\bZ_2$ 1-form symmetry and the $\SU(2n_f)/\bZ_2$ 0-form flavor symmetry form a nontrivial 2-group or not. 
This can be found rather physically by studying the line operators.

\subsection{$\Spin$}
We start by discussing the $\Spin(2n_c)$ gauge theories. 
First, recall that the center of $\Spin(2n_c)$ is $\bZ_2\times \bZ_2$ or $\bZ_4$ depending on whether $n_c$ is even or odd.
This corresponds to the fact that the tensor square of a spinor representation contains the identity representation when $n_c$ is even while it contains the vector representation when $n_c$ is odd.

We now consider the Wilson line $W$ in the spinor representation in the $\Spin(2n_c)$ gauge theory with $2n_f$ fermions in the vector representation.
When $n_c$ is even, $W^2$ contains the identity representation, and therefore we simply have a $\bZ_2$ 1-form symmetry independent of the flavor symmetry, and there is nothing to see here.

When $n_c$ is odd, $W^2$ contains the vector representation.
This can be screened by the dynamical fermion, which however carries the fundamental representation of $\su(2n_f)$ flavor symmetry, 
and in particular transforms nontrivially under $-1\in \SU(2n_f)$.
In other words, the flavor Wilson line in the vector representation of $\SU(2n_f)$ can now be considered as the square  of the gauge Wilson line in the spinor representation of $\Spin(2n_c)$.
This means that we have the following extension of groups \begin{equation}
0\to \underbrace{\bZ_2}_{\substack{\text{subgroup of}\\
\text{rep.~of center of}\\
\text{flavor $\SU(2n_f)$}}}
\to \bZ_4 
\to \underbrace{\bZ_2}_{\substack{\text{group of}\\\text{gauge Wilson lines }\\
\text{up to screening}}} \to 0.
\label{charge-extension}
\end{equation}
As the groups of charges of $\su(2n_f)$ 0-form symmetry and $\bZ_2$ 1-form symmetry are combined nontrivially, 
the symmetry groups themselves are also combined nontrivially.
This can be seen most clearly by considering their background fields.

The fermion fields are simultaneously in the vector representation of $\SO(2n_c)$ and the fundamental representation of $\SU(2n_f)$,
and therefore are in a representation of $G=[\SO(2n_c)\times \SU(2n_f)]/\bZ_2$.
Given a $G$-bundle on a manifold $X$,
there is an $\SO(2n_c)/\bZ_2$ bundle and an $\SU(2n_f)/\bZ_2$ bundle associated to it.
Let us denote by $v_2,a_2\in H^2(X;\bZ_2)$ the obstruction classes controlling whether they lift to $\SO(2n_c)$ and $\SU(2n_f)$ respectively. 
Then we have $v_2=a_2$ for a $G$-bundle.
The flavor Wilson line in the fundamental representation is charged under $-1\in \SU(2n_f)$ in the center,
and $a_2$ can be considered as the background field for this $\bZ_2$ 1-form center symmetry.

Now, without the flavor background,  the background $E\in H^2(X;\bZ_2)$ for the electric $\bZ_2$ 1-form symmetry of the $\Spin(2n_c)$ theory sets the Stiefel-Whitney class $w_2\in H^2(X;\bZ_2)$ of the $\SO(2n_c)$ gauge bundle to be $E=w_2$, which controls whether it lifts to a $\Spin(2n_c)$ bundle.
When the flavor background $a_2$ is nontrivial,
the obstruction class $v_2$ controlling the lift from $\SO(2n_c)/\bZ_2$ to $\SO(2n_c)$ is nontrivial.
In this situation when $n_c$ is odd, $w_2$ can no longer be defined as a closed cochain; rather it satisfies $\delta w_2 = \beta v_2$, where $\beta$ is the Bockstein operation,
since together they specify the obstruction class $x_2\in H^2(X;\bZ_4)$ 
controlling the lift from $\SO(2n_c)/\bZ_2=\Spin(2n_c)/\bZ_4$ to $\Spin(2n_c)$.
As $E=w_2$ and $v_2=a_2$, we conclude that the background field satisfies \begin{equation}
\delta E=\beta a_2.
\label{ba}
\end{equation}
This means that the $\bZ_2$ 1-form symmetry and the $\SU(2n_f)/\bZ_2$ 0-form flavor symmetry form the 2-group $H$ fitting in the sequence \begin{equation}
0\to \bZ_2[1]\to H \to \SU(2n_f)/\bZ_2 \to 0
\label{group-extension}
\end{equation} with the Postnikov class being $\beta a_2 \in H^3(B(\SU(2n_f)/\bZ_2);\bZ_2)$.

Before proceeding, we note that having the extension of groups of charges of line operators as in \eqref{charge-extension}
is equivalent to having a nontrivial 2-group extension \eqref{group-extension} 
whose background field satisfies \eqref{ba}.
Therefore, to find a nontrivial 2-group extension, we can simply study the group of charges of line operators,
which we will carry out for $\SO_\pm$ gauge theories next.

@@

We would like to study how the magnetic $\bZ_2$ 1-form symmetry of the $\SO(2n_c)_+$ theory is combined with the $\su(2n_f)$ flavor symmetry.
In accord with the discussions in the previous subsection, 
we consider what happens when we take two copies of the 't Hooft line operator $H$ and fuse them.
At the very naive level, $H^2$ can be screened by dynamical monopoles,
but dynamical monopoles can receive flavor/gauge center charges from the fermion zero modes.
To study these issues, it is useful to deform the theory and make it simpler
by performing the following steps:
\begin{itemize}
\item Reduce the flavor symmetry from $\su(2n_f)$ to $\usp(2n_f)$. 
The fundamental representation still transforms nontrivially under $-1\in \USp(2n_f)$, which is enough for our purposes.
\item Add an adjoint scalar $\Phi_{[ab]}$ and the interaction $\psi^{ai}_\alpha \psi^{bj}_\beta J_{ij} \Phi_{ab}\epsilon^{\alpha\beta} +c.c.$, where $J_{[ij]}$ is the constant matrix for the $\usp(2n_f)$ part.
\item Give a vacuum expectation value (vev) to $\Phi_{ab}$ to break $\SO(2n_c)$ to $\SO(2)^{n_c}$.
\end{itemize}

The 't Hooft lines in the resulting $\SO(2)^{n_c}$ theory can be labeled by their magnetic charges $(m_1,\ldots,m_{n_c})\in \bZ^{n_c}$.
The dynamical monopoles have the charges in the `adjoint class', which are
in the root lattice $\Lambda$ of $\SO(2n_c)$. 
Then, the group of the magnetic charges of 't Hooft lines up to screening by the dynamical monopoles is \begin{equation}
\bZ^{n_c}/\Lambda = \bZ_2,
\end{equation}
which agrees with the 1-form symmetry before the deformation.
We now would like to study how this $\bZ_2$ is combined with the flavor/gauge center $\bZ_2$ charge.

@@

Let us say we chose the standard $\phi$ such that the simple roots are $(1,-1,\ldots,0)$,
$(0,1,-1,\ldots,0)$, \ldots, $(0,\ldots,1,-1)$ and $(0,\ldots,1,+1)\in \bZ^{n_c}$,
which we call simple dynamical monopoles.
Now, consider the group $\bZ^{n_c}\times \bZ_2$ which combines the magnetic charges in $\bZ^{n_c}$ and the flavor/gauge center charge $q \in \bZ_2$.
What we are after is the quotient of $\bZ^{n_c}\times \bZ_2$ 
by the subgroup generated by the charges of simple dynamical monopoles,
which we denote by
$(1,-1,\ldots, 0; q_1)$, $(0,1,-1,\ldots, 0; q_2)$, \ldots,
$(0,\ldots, 1,-1; q_{n_c-1})$ and $(0,\ldots,1,+1; q_{n_c})$, respectively.

To determine this quotient,
we do not have to determine the all $q_i$'s; 
we simply use $q_{i=1,\ldots, n_c-2}$ to rewrite any charge vector $
(m_1,\ldots, m_{n_c-2}, m_{n_c-1},m_{n_c}; q)
$
into the form $
(0,\ldots,0, m, m' ; q').
$
Then, only $q_{n_c-1}$ and $q_{n_c}$ need to be determined.
This reduces the study to the case of $n_c=2$ and $\so(4)\simeq \su(2)_1 \times \su(2)_2$,
where the monopoles associated to the simple roots are just 't Hooft-Polyakov monopoles associated to the two factors of $\su(2)$'s.
The vev of the adjoint scalar in this basis can be written as $(a_1,a_2)$, which we assume to be $a_1>a_2>0$.
Here, the fermion is in the vector representation of $\so(4)$.
Under the monopole in $\su(2)_1$, it is a doublet coupled to an adjoint vev of size $a_1$ with bare mass $a_2$,
and similarly for the monopole in $\su(2)_2$.

Now, the explicit analysis in \cite[Sec.~IV]{Callias:1977kg} concerning the number of zero modes in the 't Hooft-Polyakov monopole says that 
a doublet coupled to an adjoint vev of size $a$ with bare mass $\mu$ has
a zero mode if $|a|>|\mu|$ and 
has no zero modes if $|a|<|\mu|$.
With our assumption $a_1>a_2>0$, this means that the monopole in $\su(2)_1$ has a zero mode, while the monopole in $\su(2)_2$ does not.
In our original basis, this means that the monopole with $(0,\ldots,1,-1;q_{n_c-1})$ does not produce any zero modes and $q_{n_c-1}=0$,
while the monopole with $(0,\ldots,1,+1;q_{n_c})$  has two zero modes per flavor.
The 1-form symmetry group is obtained by dividing $\bZ^2\times \bZ_2$ by the subgroup generated by $(1,-1;0)$ and $(1,+1;q_{n_c})$.
This is $\bZ_2\times \bZ_2$ or $\bZ_4$ depending on whether $q_{n_c}$ is $0$ or $1$.

Let us determine $q_{n_c}$.
We saw that there are two zero modes per flavor; 
this means that there are fermionic zero modes transforming in \begin{equation}
V_{2} \otimes R_{2n_f} 
\end{equation} where $V_2$ is the doublet of $\su(2)_2$\footnote{
	It is actually broken to $\u(1)$, but keeping $\su(2)_2$ representation is useful in organizing the answer.
}, and
$R_{2n_f}$ is the fundamental of $\usp(2n_f)$,
and we need to impose the reality condition using the pseudo-reality of both factors,
so that there are $4n_f$ Majorana fermion in total.

To determine the flavor/gauge center charge $q_{n_c}$ of the monopole,
it suffices to consider the case $n_f=1$;
the general case is given by simply multiplying it by $n_f$.
When $n_f=1$, there are $4$ Majorana fermions.
Quantizing them, we find the monopoles in \begin{equation}
(V_2 \otimes \mathbf{1}) \oplus (\mathbf{1}\otimes R_2).
\end{equation}
It has the `vector' charge under $\usp(2)\simeq\su(2)$ flavor symmetry or is a doublet under $\su(2)_2$,
which corresponds to the `vector' charge under $\so(4)$ gauge symmetry.
In either case, they have the flavor/gauge center charge $1 \in \{0,1\}=\bZ_2$.
Therefore we conclude the flavor/gauge center charge $q_{n_c}$ is simply given by $n_f$ mod 2.

@@

The analysis of the $\SO(2n_c)_-$ theory is largely the same;
the only difference is that the discrete theta angle gives an additional gauge center charge to the simple dynamical monopole with the magnetic charge $(0,0,\ldots,1,+1)$, so that $q_{n_c}=n_f+n_c$ mod 2.
Therefore, we conclude the following:
for the $\SO(2n_c)_-$ theory,
the group $\bZ_2$ of magnetic charges of 't Hooft lines is extended by the flavor/gauge center symmetry $\bZ_2$ to become $\bZ_4$ when $n_f+n_c$ is odd,
while they remain separate when $n_f+n_c$ is even.

@@

\begin{table}
\centering
\renewcommand{\arraystretch}{1.2}
\begin{tabular}{rl|ccc}
$(n_c,$&$n_f)$ & $\Spin$ & $\SO_+$ & $\SO_-$\\
\hline
(even,&even) & product & product & product \\
(odd,&even) & extended & product & extended \\
(even,&odd) & product & extended & extended \\
(odd,&odd) & extended & extended & product
\end{tabular}
\caption{How the $\bZ_2$ 1-form symmetry and the flavor symmetry $\SU(N_f)/\bZ_2$ are combined
in $\so(2n_c)$ QCD.
`product' means that they form a direct product,
and `extended' means that they form a nontrivial 2-group.
\label{table:2group}}
\end{table}

@@

Let us say that  a four-dimensional spin QFT $Q$ with $\bZ_2$ 1-form symmetry is given.
We denote its partition function on a manifold $X$ by $Z_Q[E]$, 
where we suppress the dependence on $X$ in the notation and $E\in H^2(X;\bZ_2)$ is the background field for the $\bZ_2$ 1-form symmetry.
We then define $SQ$ and $TQ$ to be QFTs with partition functions given by the formula \begin{equation}
Z_{SQ}[B] \propto \sum_{E} (-1)^{\int_X B\cup E} Z_Q[E],\qquad
Z_{TQ}[E]= (-1)^{\int_X \tfrac12 \fP(E)} Z_Q[E].
\end{equation}
We can show that $S^2=T^2=1$ and $(ST)^3=1$, meaning that they generate $\SL(2,\bZ_2)$.
This operation was considered in  \cite{Gaiotto:2014kfa} as an analogue of the $\SL(2,\bZ)$ action on 3d theories with $\U(1)$ symmetry of  \cite{Witten:2003ya} and then further studied in \cite{Bhardwaj:2020ymp}.
